Replicating Blair and Sambanis: Forecasting Civil Wars - JCR 2020

### 2020-04-28 issues.Rmd
some changes

While reading Blair and Sambanis (2020), a few issues stuck out. We downloaded the replication files authors provide from https://journals.sagepub.com/doi/suppl/10.1177/0022002720918923/suppl_file/sj-zip-1-jcr-10.1177_0022002720918923.zip. The original folder name of is `sj-zip-1-jcr-10.1177_0022002720918923/replication-3`. For simplicity, we extracted this `replication-3` and saved to the `Replication Files` folder of this repo.

@@

1. Improperly tuning of models
  - RM: In a twitter DM, I asked Blair to clarify -- "[I] was wondering how you all came to your tuning procedure for the RF model hyper-parameters?" He responded "mostly trial and error, honestly. ... Trial and error was on early data. Forecasts were for much later data"
2. Using a regression framework in a classification problem
3. Their train/validate/test approach
4. Rounding of AUC/ROC scores
5. Lack of yearly test forecasts in favor of a single 5-year test forecast
6. Table 4! 
    - There are only 20 onsets from 2001 to 2015 But there are 15 "Persistence" cases (top of table 4) and 16 "Change" cases (bottom of table 4)
    - Why only top 30?
7. Treating the output of a RF regression model as Pr() -- (RM: I guess this is akin to a Linear Probability Model, but I'm not sure if this is possible in a RF framework
8. The lack of procedures to account for rare events in an RF model
9.

>>>>

While reading Blair and Sambanis (2020), a few issues stuck out. We downloaded the replication files authors provide from https://journals.sagepub.com/doi/suppl/10.1177/0022002720918923/suppl_file/sj-zip-1-jcr-10.1177_0022002720918923.zip. The original folder name of is `sj-zip-1-jcr-10.1177_0022002720918923/replication-3`. For simplicity, we extracted this `replication-3` folder and saved to the `Replication Files` folder of this repo.

@@

- Improperly tuning of models
  - RM: In a twitter DM, I asked Blair to clarify -- "[I] was wondering how you all came to your tuning procedure for the RF model hyper-parameters?" He responded "mostly trial and error, honestly. ... Trial and error was on early data. Forecasts were for much later data"
- Using a regression framework in a classification problem
- Their train/test approach 
  - 5-year forecasts for sep plots and AUC scores)
- Rounding of AUC/ROC scores
- Lack of yearly test forecasts in favor of a single 5-year test forecast
- Table 4! 
  - There are only 20 onsets from 2001 to 2015 But there are 15 "Persistence" cases (top of table 4) and 16 "Change" cases (bottom of table 4)
- Treating the output of a RF regression model as Pr() -- (RM: I guess this is akin to a Linear Probability Model, but I'm not sure if this is possible in a RF framework
- The lack of procedures to account for rare events in an RF model

To-do

1. Full replication
    - Can we get their results
    - RM: maybe I can go through their code and bring it "up-to-date" i.e. make it tidy... 
    - Figure out how they built table 4 (It looks like this is done in STATA)
    - Check to see if it matter that they are using regression for a classification problem
2. Address tuning 
    - Does properly tuning the CAMEO model improve its performance?
3. Address AUC/ROC smoothing... 
    - How drastic is the change; does it change their conclusions? 
4. Run a yearly test forecast -- as is the current standard in TSCS forecasts (Is it the current standard?) 
5.
